---
title: "Problem_set:exp_1_var"
output: html_document
---
## Loading the data 
```{r}
library(ggplot2) #must load the ggplot package first
data(diamonds) #loads the diamonds data set since it comes with the ggplot package

#the diamonds data set will be listed as a 'Promise' in the workspace. This is a special object in R, and you need to run a command on the data to fully load the data set
summary(diamonds)
#Find out how many variables & observations in the set
dim(diamonds)
#More info on variables
str(diamonds)
?diamonds

diamonds$color
```

## Histogram of price of diamonds
```{r}
qplot(x=price, data = diamonds, color=I('black'), fill=I('#5760AB'))
```

```{r}
summary(diamonds$price)
```
* The graph is a long tailed positively skewed graph
* The mean is $3933
* The median is $2401

## Diamond counts
```{r}
#How many diamonds cost less than $500?
sum(diamonds$price<500)

#How many diamonds cost less than $250?
sum(diamonds$price<250)

#How many diamonds cost $15,000 or more?
sum(diamonds$price>=15000)
```

## Exploring the price histogram
Explore the largest peak in the price histogram you created earlier.
Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis.
```{r}
qplot(x=price, data = diamonds, binwidth=100, color=I('black'), fill=I('#5760AB')) +
  scale_x_continuous(limits=c(250,1700), breaks = seq(250, 1700, 100))
```
The large peak on the original graph occurs witin the region of 0 - 1000. Limit the x axis to 1500 in order to get a better look at this region.
* We see the majority of diamonds in this region are priced $650 - $750
* There are barely any diamonds priced less than $350
* There is a significant gap at the $1450 - $1550 range


## Break out the histogram of diamond prices by cut. 

```{r}
qplot(x=price, data=diamonds, color=I('black'), fill=I('#5760AB')) +
  facet_wrap(~cut)
by(diamonds$price, diamonds$cut, summary)
```
If we look at the output table, the median and quartiles are reasonably close to each other. This means the distributions should be somewhat similar, but the histograms we created don't show that.
The 'Fair' and 'Good' diamonds appear to have different distributions compared to the better cut diamonds. They seem somewhat uniform on the left with long tails on the right.
We will add the *free_y* parameter to facet_wrap so that the y-axis in the histograms is not fixed.
```{r}
?facet_wrap
qplot(x=price, data=diamonds, color=I('black'), fill=I('#5760AB')) +
  facet_wrap(~cut, scales = "free_y")
```

## Price per carat by cut
```{r}
qplot(x=price/carat, data=diamonds, binwidth=0.25, color=I('black'), fill=I('#5760AB')) +
  scale_x_log10() +
  facet_wrap(~cut, scales = "free_y")
```


## Price box plots

Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

```{r}
ggplot(aes(x=color, y=price), data=diamonds) +
  geom_boxplot()

by(diamonds$price, diamonds$color, summary)
#Get IQR for diamonds with the best and worst colours
IQR(subset(diamonds, color=="D")$price)
IQR(subset(diamonds, color=="J")$price)
```
* Reminder: "D"<"E"<"F"<"G"<.. in terms of the best colours
* We can see that D coloured diamond which was supposed to be the best quality has a lower mean prices than J coloured which was supposed to be the worst quality.
* D also had the lowest max price, second lowest median, and second lowest mean, when I expected it to have the greatest for all three measures. This suggests that diamond sellers are really not setting price by the color of a diamond. There's likely a confounding variable like the carat or clarity that matters more.
* Check to see what the price per carat is for each of the colours:
```{r}
ggplot(aes(x=color, y=price/carat), data=diamonds) +
  geom_boxplot()
```
There's less of a discrepancy between D & J median prices and IQRs when measuring price per carat which shows price per carat doesn't vary a lot over the different colours. Therefore the carat is more likely to affect the price than colour

## carat frequency polygon
```{r}
ggplot(aes(x=carat,), data=diamonds) +
  geom_freqpoly(aes(color=carat), binwidth=0.1) +
  scale_x_continuous(lim=c(0,3), breaks = seq(0,3,0.1) ) +
  scale_y_continuous(breaks = seq(0,16000, 2000))

table(diamonds$carat)

```

## facebook friends
Questions to answer
1.How many people share your birthday? Do you know them?
2. Which month contains the most number of birthdays?
3. How many birthdays are in each month?
4. Which day of the year has the most number of birthdays?
5. Do you have at least 365 friends that have birthdays on everyday of the year?

```{r}
FB_birthdays <- read.csv('birthdaysExample.csv', se = ',')
str(FB_birthdays)
```


Once you load the data into R Studio, you can use the strptime() function
to extract the birth months and birth days.
```{r}
d1 <- strptime(FB_birthdays$dates, format='%m/%d/%y')
d2 <- strptime(FB_birthdays$dates, '%m/%d/%y')

m <- strftime(d2, format='%m')
d <- strftime(d2, format='%d')

qplot(x=d, xlab="dates")
qplot (x=m, xlab="months")

```


